---
title: "R Notebook"
output: html_notebook
---

Trying to run discrete Fourier series data imputation models using stan. 

Set up the workspace
```{r}
rm(list = ls())
library(tidyverse)
library(rstan)

library(jagsUI)
library(coda)

source("Data_Imputation_funcs.R")    

MCMC.n.chains <- 5
MCMC.n.samples <- 500000
MCMC.n.burnin <- 350000
MCMC.n.thin <- 50

n.cores <- parallel::detectCores()
MCMC.setup <- list(warmup = MCMC.n.burnin,
                   iter = MCMC.n.samples,
                   chains = ifelse(n.cores < 5, n.cores, 5))

```


Bring in the data. First, I just use JM data.

```{r}

all.years <- 2001:2017
idx <- 1:length(all.years)

year.begin.JM <- 2001
year.end <- 2017
data.JM <- data.extract(location = "JM", 
                        year.begin = year.begin.JM, 
                        year.end = year.end)

JM.keep <- 2001:2017
idx.JM <- idx[all.years %in% JM.keep]
n.keep.JM <- length(idx.JM)
dt.JM <- idx.JM[2:length(idx.JM)] - idx.JM[1:(length(idx.JM)-1)]

y <- as.vector(t(data.JM$jags.data2$y))
n.years <- 17
n.months <- 12
period_1 <- 12
c.const <-  2 * pi * c(1:12) / period_1

data_list <- list(n_months = 12, 
                  n_years = n.years,
                  years = rep(2001:2017, each = 12),
                  period_1 = 12, 
                  m = rep(1:12, times = n.years),
                  N_obs = sum(!is.na(y)),
                  N_miss = sum(is.na(y)),
                  idx_obs = which(!is.na(y)),
                  idx_miss = which(is.na(y)),
                  y_obs = y[which(!is.na(y))],
                  y_miss = y[which(is.na(y))],
                  c_const = c.const)

```


Then run stan - it's just too slow... 
# ```{r}
# 
# if (!file.exists("RData/DFS_norm_norm_stan.rds")){
# 
#   mod.out.stan <- stan("models/Model_DFS_norm_norm.stan", 
#                        data=data_list, 
#                        chains = MCMC.setup$chains, 
#                        warmup = MCMC.setup$warmup,
#                        iter = MCMC.setup$iter,
#                        cores = getOption("mc.cores", MCMC.setup$chains),
#                        control = list(adapt_delta = 0.999,
#                                       max_treedepth = 12))
#   saveRDS(mod.out.stan, file = "RData/DFS_norm_norm_stan.rds")
#   
# } else {
#   mod.out.stan <- readRDS("RData/DFS_norm_norm_stan.rds")
# }
# ```


<!-- Didn't converge...  -->
<!-- ```{r} -->
<!-- traceplot(mod.out.stan, pars = c("beta_sin", "beta_cos")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- traceplot(mod.out.stan, pars = c("sigma_X", "sigma_y")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- traceplot(mod.out.stan, pars = "predX0") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- traceplot(mod.out.stan, pars = c("y_miss[1]", "y_miss[2]")) -->
<!-- ``` -->

<!-- Extract the log-likelihood and run the loo. -->
<!-- ```{r} -->
<!-- log_lik <- extract_log_lik(mod.stan,  -->
<!--                            merge_chains = FALSE) -->
<!-- rel_n_eff <- relative_eff(exp(log_lik)) -->
<!-- loo(log_lik, r_eff = rel_n_eff, cores = 4) -->
<!-- ``` -->


Try the exactly the same model in JAGS:

```{r}

MCMC.params <- list(n.chains = MCMC.n.chains,
                    n.samples = MCMC.n.samples,
                    n.burnin = MCMC.n.burnin,
                    n.thin = MCMC.n.thin)

jags.params <- c("predX0", "X", "y", "loglik",
                 "sigma.y", "sigma.X", "beta_cos", "beta_sin")

data.jags <- data.JM$jags.data
data.jags$c_const = c.const

#model.file = 'models/model_SSAR1_logY_norm_norm_var.txt'
if (!file.exists("RData/DFS_norm_norm_JAGS.rds")){
  
  mod.out.jags <- jagsUI(data.jags, 
                         parameters.to.save = jags.params, 
                         model.file = "models/model_DFS_norm_norm_JAGS.txt", 
                         n.chains = MCMC.n.chains,
                         n.iter = MCMC.n.samples,
                         n.burnin = MCMC.n.burnin,
                         n.thin = MCMC.n.thin,
                         parallel = T)
  
  saveRDS(mod.out.jags, "RData/DFS_norm_norm_JAGS.rds")
  
} else {
  mod.out.jags <- readRDS("RData/DFS_norm_norm_JAGS.rds")
}
```

A LOT faster than stan!

```{r}
bayesplot::mcmc_trace(mod.out.jags$samples, pars = c("beta_sin", "beta_cos"))


```
```{r}
bayesplot::mcmc_trace(mod.out.jags$samples, pars = c("sigma.X", "sigma.y"))
```

```{r}
bayesplot::mcmc_trace(mod.out.jags$samples, pars = "predX0")
```

```{r}
idx.y.NA <- which(is.na(data.jags$y))

idx.y.NA.first.4 <- idx.y.NA[1:4]

# can't figure out how to dynamically provide the pars here... I've done this before
# bayesplot::mcmc_trace(mod.out.jags$samples, 
#                       pars = c(paste0("\"y[", idx.y.NA.first.4[1], "]\",",
#                                       "\"y[", idx.y.NA.first.4[2], "]\",",
#                                       "\"y[", idx.y.NA.first.4[3], "]\",",
#                                       "\"y[", idx.y.NA.first.4[4], "]\"")))

bayesplot::mcmc_trace(mod.out.jags$samples,
                      pars = c("y[8]", "y[9]", "y[10]", "y[11]"))

```

Looks pretty good to me! Not sure why Stan takes so long to do this... 

Do Pareto K GOF analysis:
```{r}

n.per.chain <- (mod.out.jags$mcmc.info$n.iter - mod.out.jags$mcmc.info$n.burnin)/mod.out.jags$mcmc.info$n.thin
#   
# log likleihood values (n.per.chain x n.chains) x # data points
loglik.norm.norm <- mod.out.jags$sims.list$loglik

# convert the data (y) into a vector also:
y <- data.jags$y

# get rid of NA columns - even if data existed (for example the first value) - no likelihood
# for the first data point - this is not an issue for this case so skip
loglik.norm.norm.obs <- loglik.norm.norm[, !is.na(y)]

Reff.norm.norm <- loo::relative_eff(exp(loglik.norm.norm.obs),
                                    chain_id = rep(1:mod.out.jags$mcmc.info$n.chains,
                                                   each = n.per.chain),
                                    cores = 5)
#   
loo.out.norm.norm <- loo(loglik.norm.norm.obs, r_eff = Reff.norm.norm, cores = 5)
#   ####
# loo.out <- pareto.k.diag.3D(jm, MCMC.params, jags.data)

```

plot model fit statistics - pareto k - they look all good. 

```{r}

pareto.k.norm.norm <- loo.out.norm.norm$diagnostics$pareto_k
data.y <- na.omit(y)
# 
norm.norm.pareto.df <- data.frame(y = data.y,
                                  khat = pareto.k.norm.norm,
                                  datapoint = seq(from = 1, to = length(data.y)),
                                  k0.7 = cut(pareto.k.norm.norm,
                                             breaks = c(0, 0.7, 1.5),
                                             labels = c("<=0.7", ">0.7")))
p.2 <- ggplot(data = norm.norm.pareto.df) +
  geom_path(aes(x = datapoint, y = y), alpha = 0.5) +
  geom_point(aes(x = datapoint, y = y,
                 size = khat,
                 color = k0.7)) +
  scale_size_continuous(limits = c(0.0, 1.3),
                        range = c(1, 4))+
  scale_color_manual(values = c("<=0.7" = "darkgreen",
                                ">0.7" = "red"))
p.2
```

Try Student's t observation model to add more uncertainty to the observations - start here 2020-06-29

The jags code needs to be completed as well. Parameters are not set yet. 

```{r}
jags.params <- c("predX0", "X", "y", "loglik", "df",
                 "sigma.y", "sigma.X", "beta_cos", "beta_sin")

data.jags <- data.JM$jags.data
data.jags$c_const = c.const

#model.file = 'models/model_SSAR1_logY_norm_norm_var.txt'
if (!file.exists("RData/DFS_norm_t_JAGS.rds")){
  
  mod.out.norm.t.jags <- jagsUI(data.jags, 
                         parameters.to.save = jags.params, 
                         model.file = "models/model_DFS_norm_t_JAGS.txt", 
                         n.chains = MCMC.n.chains,
                         n.iter = MCMC.n.samples,
                         n.burnin = MCMC.n.burnin,
                         n.thin = MCMC.n.thin,
                         parallel = T)
  
  saveRDS(mod.out.norm.t.jags, "RData/DFS_norm_t_JAGS.rds")
  
} else {
  mod.out.norm.t.jags <- readRDS("RData/DFS_norm_t_JAGS.rds")
}
```

Do Pareto K GOF analysis:
```{r}

n.per.chain <- (mod.out.norm.t.jags$mcmc.info$n.iter -
                  mod.out.norm.t.jags$mcmc.info$n.burnin)/mod.out.norm.t.jags$mcmc.info$n.thin
#   
# log likleihood values (n.per.chain x n.chains) x # data points
loglik.norm.t <- mod.out.norm.t.jags$sims.list$loglik

# convert the data (y) into a vector also:
y <- data.jags$y

# get rid of NA columns - even if data existed (for example the first value) - no likelihood
# for the first data point - this is not an issue for this case so skip
loglik.norm.t.obs <- loglik.norm.t[, !is.na(y)]

Reff.norm.t <- loo::relative_eff(exp(loglik.norm.t.obs),
                                 chain_id = rep(1:mod.out.norm.t.jags$mcmc.info$n.chains,
                                                each = n.per.chain),
                                 cores = 5)
#   
loo.out.norm.t <- loo(loglik.norm.t.obs, r_eff = Reff.norm.t, cores = 5)
#   ####
# loo.out <- pareto.k.diag.3D(jm, MCMC.params, jags.data)

```

All Pareto k diagnostic values are bad... so we should go with the norm-norm model.

